### 2018-05-22 inst/doc/command-line-example.Rmd

---
title: "LDATS Example"
author: "Juniper L. Simonis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include=FALSE}
library(LDATS)
vers <- packageVersion("LDATS")
today <- Sys.Date()
```

This vignette walks through an example of **LDATS** at the command line and
was constructed using **LDATS** version `r vers` on `r today`.

## Installation

To obtain the most recent version of **LDATS**, install the most recent 
version from GitHub:

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("weecology/LDATS")
```

## Data

For this vignette, we will be using rodent data from the control plots of the
[Portal Project](https://github.com/weecology/portaldata), which come with
the **LDATS** package

```{r}
data(rodents)
rodents
```