### 2018-06-22 previous_work/paper/markdown_analyses.Rmd
updating the previous work folder

---
title: "LDA kratplots results"
author: "Renata Diaz"
date: "5/8/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(topicmodels)
library(RCurl)
library(multipanelfigure)
library(reshape2)
library(LDATS)

source('rodent_data_for_LDA.r')
source('rodent_data_from_portalr.R')
source('AIC_model_selection.R')
source('LDA_figure_scripts.R')
source('changepointmodel.r')
source('LDA-distance.R')

```

## Eight plots, starting in 1988

### Prepare rodent data

```{r}

dat = get_exclosure_rodents(cont_or_exclosures = 'control', time_or_plots = 'plots')

# dates to go with count data
moondat = read.csv(text=getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"),stringsAsFactors = F)
moondat$date = as.Date(moondat$censusdate)

period_dates = filter(moondat,period %in% dat$period) %>% select(period,date)
dates = period_dates$date

dat = dat[,2:22]

```

### Select number of topics
```{r}
# ==================================================================
# 2a. select number of topics
# ==================================================================
Select LDA model using LDATS

# Changepoint models

# set up parameters for model
year_continuous = 1970 + as.integer(julian(dates)) / 365.25
x = data.frame(
  year_continuous = year_continuous,
  sin_year = sin(year_continuous * 2 * pi),
  cos_year = cos(year_continuous * 2 * pi)
)

# run models with 1, 2, 3, 4, 5 changepoints
cp_results_rodent_3topics = changepoint_model(ldamodel3, x, 1, weights = rep(1,length(year_continuous)))
cp_results_rodent2_3topics = changepoint_model(ldamodel3, x, 2, weights = rep(1,length(year_continuous)))
cp_results_rodent3_3topics = changepoint_model(ldamodel3, x, 3, weights = rep(1,length(year_continuous)))
cp_results_rodent4_3topics = changepoint_model(ldamodel3, x, 4, weights = rep(1,length(year_continuous)))
cp_results_rodent5_3topics = changepoint_model(ldamodel3, x, 5, weights = rep(1,length(year_continuous)))

# Save changepoint models
save(cp_results_rodent_3topics,
     cp_results_rodent2_3topics, 
     cp_results_rodent3_3topics, cp_results_rodent4_3topics, 
     cp_results_rodent5_3topics, 
     file = 'models/cp_8plots_3topics.Rdata')

# change point model selection
# mean deviance ( -2 * log likelihood) + 2*(#parameters)
mean(cp_results_rodent_3topics$saved_lls * -2) + 2*(3*(ntopics-1)*(1+1)+(1))
mean(cp_results_rodent2_3topics$saved_lls * -2)+ 2*(3*(ntopics-1)*(2+1)+(2))
mean(cp_results_rodent3_3topics$saved_lls * -2)+ 2*(3*(ntopics-1)*(3+1)+(3))
mean(cp_results_rodent4_3topics$saved_lls * -2)+ 2*(3*(ntopics-1)*(4+1)+(4))
mean(cp_results_rodent5_3topics$saved_lls * -2)+ 2*(3*(ntopics-1)*(5+1)+(5))

# 3 topics is best supported.

rm(list= c(cp_results_rodent2_3topics, cp_results_rodent_3topics,
           cp_results_rodent4_3topics, cp_results_rodent5_3topics))

# some quick histograms of changepoint model results
hist(year_continuous[cp_results_rodent3_3topics$saved[,1,]],breaks = seq(1988,2016,.25),xlab='',main='Changepoint Estimate')
annual_hist(cp_results_rodent3_3topics,year_continuous)

# turn changepoint results into data frame
df_3_3topics = as.data.frame(t(cp_results_rodent3_3topics$saved[,1,])) %>% melt()
df_3_3topics$value = year_continuous[df_3_3topics$value]

# find 95% confidence intervals on each changepoint:
quantile(df_3_3topics[df_3_3topics$variable=='V1','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
quantile(df_3_3topics[df_3_3topics$variable=='V2','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')
quantile(df_3_3topics[df_3_3topics$variable=='V3','value'],probs=c(.025,.975)) %>% date_decimal() %>% format('%d-%m-%Y')

write.csv(df_3_3topics, file = 'models/df_3_3topics.csv', row.names =F)

```

#### Eight plots starting in 1988, 4 topics
```{r}
SEED_4 = unlist(best_seed_4)[1]

ntopics = 4

ldamodel4 = LDA(dat,ntopics, control = list(seed = SEED_4),method='VEM')

```