### 2018-10-26 inst/doc/LDATS.Rmd

---
title: "Latent Dirichlet Allocation Time Series (LDATS)"
author: "Juniper L. Simonis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include=FALSE}
library(LDATS)
vers <- packageVersion("LDATS")
today <- Sys.Date()
```

## Overview

This vignette details the code base for the **LDATS** package. It was
constructed using **LDATS** version `r vers` on `r today`.

## Installation

To obtain the most recent version of **LDATS**, install the most recent 
version from GitHub:

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("weecology/LDATS")
```

## Analytical Pipeline

* `LDA_TS()`
  * `LDA_set()`
    * `LDA()`
  * `select_LDA()`
  * `TS_set_on_LDA()`
  * `select_TS()`