### 2018-12-13 rmd-ignore/compare-christensen.Rmd
rmd-ignore folder contains comparison w ecology paper

---
title: "Comparison with Christensen paper"
author: "Renata Diaz"
date: "12/13/2018"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Side-by-side comparison of LDATS results (and Portal rodents vignette) with analysis from Christensen et al 2018.

## LDATS Installation

To obtain the most recent version of **LDATS**, install the most recent 
version from GitHub:

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("weecology/LDATS")
```

```{r, eval = T}
library(LDATS)
```

## Christensen 2018 analysis

Download Christensen 2018 analysis scripts & data files from [Extreme-events-LDA repo:](https://github.com/emchristensen/Extreme-events-LDA)

Main Analysis Scripts:

  * rodent_LDA_analysis.R main script for analyzing rodent community change using LDA

  * rodent_data_for_LDA.R contains a function that creates the rodent data table used in analyses

  * AIC_model_selection.R contains functions for calculating AIC for different candidate LDA models

  * changepointmodel.r contains change-point model code

  * LDA-distance.R function for computing Hellinger distance analyses

Data:

  * Rodent_table_dat.csv table of rodent data, created by rodent_data_for_LDA.R

Figure scripts:

  * LDA_figure_scripts.R contains functions for making main plots in manuscript (Fig 1). Called from rodent_LDA_analysis.R

```{r download Christensen, eval = FALSE}

files_to_download <- c('rodent_LDA_analysis.r', 'rodent_data_for_LDA.r', 'AIC_model_selection.R', 'changepointmodel.r', 'LDA-distance.R', 'Rodent_table_dat.csv', 'LDA_figure_scripts.R')

for(i in 1:length(files_to_download)) {
  download.file(url = paste0("https://raw.githubusercontent.com/emchristensen/Extreme-events-LDA/master/", files_to_download[i]),
                destfile = paste0('christensen-ecology/', files_to_download[i]))
}

rm(files_to_download)
rm(i)

```

## Data

The Portal rodents control data is included in the LDATS package:

```{r LDATS data}

data(rodents)

ldats_dat <- rodents[[1]]

head(ldats_dat)

```

Load the data used in Christensen et al: