### 2020-04-21 doc/LDATS_codebase.Rmd
codebase vignette
updated to the current codebase

---
title: "Latent Dirichlet Allocation Time Series (LDATS)"
author: "Juniper L. Simonis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{LDATScodebase}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include=FALSE}
library(LDATS)
vers <- packageVersion("LDATS")
today <- Sys.Date()
```

## Overview

This vignette outlines the code base for the **LDATS** package. 
It was constructed using **LDATS** version `r vers` on `r today`.

## Installation

To obtain the most recent version of **LDATS**, install the most recent version from GitHub:

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("weecology/LDATS")
```

## Analytical Pipeline

The full LDATS analysis can be executed using the `LDA_TS` function, which is the top-level of the coding pipeline's hierarchy of functions:

* `LDA_TS()`
  * `LDA_TS_control()`
    * `LDA_control()`
    * `TS_control()`
  * `LDA()`
    * `prepare_LDA()`
      * `check_LDA()`
      * `LDA_control()`
      * `conform_data()`
    * `run_LDA()`
      * `LDA_call()` (replicated for each model)
        * `LDA_msg()`
        * `soft_call()` using `LDA$control$model` with `LDA$control$model_args`
          * `topicmodels::LDA()` (*default*) 
      * `package_LDA()`
        * `select_LDA()`
          * `soft_call()` using `LDA$control$selector with LDA$control$selector_args
            * `min()` (*default*) 
            * `measure_LDA()` (replicated for each model)
              * `soft_call()` using `LDA$control$measurer` with `LDA$control$measurer_args`
            * `AIC()` (*default*)

  * `TS()`
    * `prepare_TS()`
      * `TS_control()`
      * `run_TS()`
        * `TS_call()` (replicated for each model)
          * `TS_msg()`
          * `TS$control$model with TS$control$model_args
            * `sequential_TS_control()`
            * `est_changepoints()`
              * `TS$control$method` with `TS$control$method_args`
                * `ldats_classic_control()`
                  * `prep_saves()`
                  * `prep_ptMCMC_inputs()`
                    * `prep_temp_sequence()`
                    * `prep_proposal_dist()`
                  * `prep_cpts()`
                    * `prep_temp_sequence()`
                    * `TS$control$response` with `TS$control$response_args`
                  * `prep_ids()`
                  * `prep_pbar()`
                  * `update_pbar()`
                  * `step_chains()`
                    * `propose_step()`
                      * `proposed_step_mods()`
                      * `TS$control$response` with `TS$control$response_args`
                    * `eval_step()`
                    * `accept_step()`
                  * `swap_chains()`
                  * `update_saves()`
                  * `update_cpts()`
                  * `update_ids()`
                  * `process_saves()`
                    * `count_trips()`
             * `est_regressors()`
               * `TS$control$response` with `TS$control$response_args`
             * `package_sequential_TS()`
               * `summarize_rhos()`
               * `measure_rho_vcov()`
               * `summarize_eta()`
               * `measure_eta_vcov()`
       * `package_TS()`
         * `select_TS()`
           * `TS$control$selector` with `TS$control$selector_args`
           * `measure_TS()` (replicated for each model)
             * `TS$control$measurer` with `TS$control$measurer_args`
  * `package_LDA_TS()`

Each component model's function (`LDA()` and `TS()`) can be run independently, as well.

## Controls Lists

To minimize the length of argument lists and facilitate simple default usage throughout the pipeline, we implement an options/controls list approach, where each of the main functions (`LDA_TS`, `LDA`, and `TS`) and their subfunctions have a `control` argument that takes a `list` to replace the defaults returned by its `_control` function:

* `LDA_TS_control()`
* `LDA_control()`
* `TS_control()`