### 2023-09-17 doc/LDATS_codebase.Rmd

---
title: "LDATS Codebase"
author: "Juniper L. Simonis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{LDATS Codebase}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include=FALSE}
library(LDATS)
vers <- packageVersion("LDATS")
today <- Sys.Date()
```

## Overview

This vignette outlines the code base for the **LDATS** package. It was
constructed using **LDATS** version `r vers` on `r today`.

## Installation

To obtain the most recent version of **LDATS**, install the most recent 
version from GitHub:

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("weecology/LDATS")
```

## Analytical Pipeline

The full LDATS analysis can be executed using the `LDA_TS` function,
which is the top-level of the coding pipeline's hierarchy of
functions:

* `LDA_TS()`
  * `check_LDA_TS_inputs()`
    * `check_timename()`
    * `check_formulas()`
    * `check_nchangepoints()`
    * `check_weights()`
    * `check_control()`
    * `check_document_term_table()`
    * `check_document_covariate_table()`	
    * `check_topics()`
    * `check_seeds()`
  * `LDA_set()`
    * `check_LDA_TS_inputs()`
      * `check_control()`
      * `check_document_term_table()`
      * `check_topics()`
      * `check_seeds()`
    * loop over models
      * `prep_LDA_control()`
      * `LDA()`
  * `select_LDA()`
    * applies the `measurer` and `selector` functions supplied via `LDA_controls_list()`
  * `TS_on_LDA()`
    * `check_LDA_TS_inputs()`
      * `check_LDA_models()`
      * `check_formulas()`
      * `check_nchangepoints()`
      * `check_timename()`
      * `check_control()`
      * `check_weights()`
      * `check_document_covariate_table()`	
    * `expand_TS()`
    * loop over models
      * `prep_TS_data()`
      * `TS()`
        * `check_TS_inputs()`
          * `check_formulas()`
          * `check_nchangepoints()`
          * `check_control()`
          * `check_weights()`
        * `est_changepoints()`
          * `prep_saves()`
          * `prep_ptMCMC_inputs()`
          * `prep_cpts()`
            * loop over chains
              * `multinom_TS()`
          * `prep_ids()`
          * `prep_pbar()`
          * loop over iterations
            * `update_pbar()`
            * `step_chains()`
              * `propose_step()`
                * `proposed_step_mods()`
                  * loop over chains
                    * `multinom_TS()`
              * `eval_step()`
              * `take_step()`
            * `swap_chains()`
                * loop over chain neighbors
            * `update_saves()`
            * `update_cpts()`
            * `update_ids()`
        * `est_regressors()`
          * loop over unique realizations of change point locations
            * `multinom_TS()`
            * loop over chunks
              * `mirror_vcov()`
              * `rmvnorm()`
        * `summarize_TS()`
          * `diagnose_ptMCMC()`
            * `count_trips()`
          * `summarize_rhos()`
          * `measure_rho_vcov()`
          * `summarize_etas()`
          * `measure_eta_vcov()`
    * `select_TS()`
      * applies the `measurer` and `selector` functions supplied via `TS_controls_list()`
    * `package_LDA_TS()`

Each component model's function (`LDA_set()` and `TS()`) can be run
independently, as well.

## Controls Lists

To minimize the length of argument lists and facilitate simple default usage
throughout the pipeline, we implement an options/controls list approach, where
each of the main functions (`LDA_TS`, `LDA_set`, and `TS`) and its subfunctions
have a `control` argument that takes a `list` to replace the defaults returned 
by its `_control` function:

* `LDA_TS_control()`
* `LDA_set_control()`
* `TS_control()`